---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Unzip activity.csv if not present and load it. Convert date variable into proper format

``` {r}
library("plyr")
library("dplyr")
library("lubridate")
library("ggplot2")

if (!file.exists("activity.csv")) {
    unzip("activity.zip")
}

data <- tbl_df(read.csv("activity.csv"))
data$real_date <- ymd(data$date)

```
## What is mean total number of steps taken per day?

```{r}
agg <- summarise(group_by(data, real_date), step_count = sum(steps))

ggplot(agg, aes(x=real_date, y=step_count)) + 
    geom_bar(stat="identity", fill = 'darkblue') + 
    theme(axis.text.x=element_text(angle=90)) + 
    xlab("Days") + 
    ylab("Step count")

#mean and mediancount of steps per day
mean(agg$step_count, na.rm=TRUE) 
median(agg$step_count, na.rm=TRUE) 

```

## What is the average daily activity pattern?

```{r}

agg2 <- summarise(group_by(data, interval), mean_step_count = mean(steps, na.rm=TRUE))

ggplot(agg2, aes(x=interval, y=mean_step_count)) + 
     geom_line(color="darkblue") + 
     xlab("Interval") + 
     ylab("Step count (mean)")


#5 minute interval with maximum steps, averaged across all days
agg2$interval[which.max(agg2$mean_step_count)]

```


## Imputing missing values

There are NA's only in staps variable as you can see in summary below. I choose 
to impute values based on interval mean value.

As you can see below the mean value has not been affected, however median value
become equal to the mean.

```{r}
# Summary of data - number of NA should be visible for each variable
summary(data)

# let's replace NA with mean value for the specific interval
new_data <-
    data %>%
    left_join(agg2, by=c("interval" = "interval")) %>% #join with aggregate
    mutate(steps = ifelse(is.na(steps), mean_step_count, steps)) %>% #update steps if needed
    select(1:4) # get rid of last row with aggregates

# No more NA'S
summary(new_data)

# recalculate aggregates and draw histograms
agg3 <- summarise(group_by(new_data, real_date), step_count = sum(steps))

ggplot(agg3, aes(x=real_date, y=step_count)) + 
    geom_bar(stat="identity", fill = 'darkblue') + 
    theme(axis.text.x=element_text(angle=90)) + 
    xlab("Days") + 
    ylab("Step count")

#mean and mediancount of steps per day
mean(agg3$step_count, na.rm=TRUE) 
median(agg3$step_count, na.rm=TRUE) 

```
## Are there differences in activity patterns between weekdays and weekends?

```{r}
#add new factor to the dataset
new_data <- mutate(new_data, weekday = as.factor(mapvalues(wday(real_date), from = 1:7, to=c("weekend", "weekday", "weekday", "weekday", "weekday", "weekday", "weekend") )))

summary(new_data)

#this time I will aggregate on the fly
ggplot(new_data, aes(x=interval, y=steps)) + 
    stat_summary(fun.y="sum", geom="line", color="darkblue") +
     xlab("Interval") + 
     ylab("Step count (mean)") +
     facet_grid(weekday ~ .)

```
